>

### 2025-03-12 docs/article.qmd
Revise the methods section

Graph theory is a formal concept depicting connections among objects. A graph consists of nodes and edges, denoted as $\mathcal{G} = (\mathcal{V}, \mathcal{E})$. The nodes form a finite set, $\mathcal{V} = \{\mathcal{v}_1, \mathcal{v}_2, \dots, \mathcal{v}_n\}$, while the edges $\mathcal{E}$ represent the connection between these nodes. In a directed graph, the edge is an ordered pair of nodes, $(\mathcal{v}_i, \mathcal{v}_j) \in \mathcal{E}$, where the order of $i, j$ matters. Directed graphs may contain self connections, $(\mathcal{v}_i, \mathcal{v}_i) \in \mathcal{E}$. In an undirected graph, the edge is an unordered set of nodes, $\{\mathcal{v}_i, \mathcal{v}_j\} \in \mathcal{E}$, where the order of $i, j$ does not matter. Typically, undirected graphs do not contain self connections, $\{\mathcal{v}_i, \mathcal{v}_i\} \notin \mathcal{E}$ [@estrada2012structure].

A graph can be represented as an adjacency matrix $A$. The adjacency matrix $A$ has dimensions of $|\mathcal{V}| \times |\mathcal{V}|$, where $|\mathcal{V}|$ is the number of nodes in a graph. Each element of $a_{ij}$ is the edge: $(\mathcal{v}_i, \mathcal{v}_j) \in \mathcal{E}$ in a directed graph, or $\{\mathcal{v}_i, \mathcal{v}_j\} \in \mathcal{E}$ in an undirected graph. In an unweighted graph, $a_{ij} = 1$ if a connection exists between $\mathcal{v}_i$ and $\mathcal{v}_j$, and $a_{ij} = 0$ otherwise. In a weighted graph, $a_{ij}$ can take any real value ($a_{ij} \in \mathbb{R}$) depending on the application [@estrada2012structure].

@@

Centrality measures formalize the identification of important nodes in a graph [@estrada2012structure]. The centrality of a node is measured differently depending on the context of analysis. Previous publication on DPN highlighted four common centrality measures, namely degree, betweenness, closeness, and eigenvector centrality [@Miglio2021]. Degree centrality in a DPN describes the number of co-prescription with the medication of interest. High (low) degree centrality means the medication is often (seldom) co-prescribed. Betweenness centrality indicates the frequency of a medication connecting two other medications by the shortest possible path. High (low) betweenness centrality means the medication is (not) a "bridge" between different kind of medications. Closeness centrality is the average distance between one medication to all other medications in the DPN. High (low) closeness centrality means that the medication is (not) commonly co-prescribed. Eigenvector centrality reflects the number of co-prescription with medications that have vital role in the DPN. High (low) eigenvector centrality means that the medication is often (seldom) co-prescribed with other important medications.

The choice of centrality largely depends on the objective of network analysis. As a general guide, degree centrality is useful to identify popular medication and monitor drug overuse. Betweenness centrality is suitable for targeting for drug-interaction study and optimizing therapy plan. Closeness centrality indicates widely-used key medications and efficiency in treatment networks. Eigenvector centrality is helpful to identify influential medications and narrow down high-impact medications for drug monitoring. This study focused on eigenvector centrality to evaluate which medications have a significant influence on how antidepressants and anxiolytics are prescribed.

## Measuring relative importance

{{< include exec-summary/_measure-eigen.qmd >}}

@@

{{< include exec-summary/_spectral.qmd >}}

>>>>

Mathematically, a graph $\mathcal{G}$ is a pair $\mathcal{G} = (\mathcal{V}, \mathcal{E})$, where $\mathcal{V}$ is the set of nodes and $\mathcal{E}$ is the set of edges. In a DPN, the node set $\mathcal{V} = \{\mathcal{v}_1, \mathcal{v}_2, \dots, \mathcal{v}_n\}$ represents a finite set of $n$ medications. Additionally, $\mathcal{E}$ is a finite set of edges, where each edge is a pair of nodes that corresponds to a co-prescription. In an undirected graph, the order of $i, j$ does not matter. Typically, undirected graphs do not contain self connections, $\{\mathcal{v}_i, \mathcal{v}_i\} \notin \mathcal{E}$ [@estrada2012structure].

A graph can be represented as an adjacency matrix $A$. The adjacency matrix $A$ has dimensions of $|\mathcal{V}| \times |\mathcal{V}|$, where $|\mathcal{V}|$ is the number of nodes in a graph. Each element of $a_{ij}$ is the edge, where in an unweighted graph, $a_{ij} = 1$ if a connection exists between $\mathcal{v}_i$ and $\mathcal{v}_j$, and $a_{ij} = 0$ otherwise. In a weighted graph, $a_{ij}$ can take any real value ($a_{ij} \in \mathbb{R}$) depending on the application [@estrada2012structure].

@@

Centrality measures formalize the identification of important nodes in a graph [@estrada2012structure]. Different node centrality measures quantify different structural properties of a node. Previous work on DPNs highlighted four centrality measures that could be used to assess the importance of a medication within a co-prescription network [@Miglio2021]. Degree centrality in a DPN describes the number of co-prescription with the medication of interest. High (low) degree centrality means the medication is often (seldom) co-prescribed. Betweenness centrality indicates the frequency of a medication connecting two other medications by the shortest possible path. High (low) betweenness centrality means the medication is (not) a "bridge" between different kind of medications. Closeness centrality is the average distance between one medication to all other medications in the DPN. High (low) closeness centrality means that the medication is (not) commonly co-prescribed. Eigenvector centrality reflects the number of co-prescription with medications that have vital role in the DPN. High (low) eigenvector centrality means that the medication is often (seldom) co-prescribed with other important medications.

The choice of centrality largely depends on the objective of network analysis. As a general guide, degree centrality is useful to identify popular medication and monitor drug overuse. Betweenness centrality is suitable for targeting for drug-interaction study and optimizing therapy plan. Closeness centrality indicates widely-used key medications and efficiency in treatment networks. Eigenvector centrality is helpful to identify influential medications and narrow down high-impact medications for drug monitoring. This study focused on eigenvector centrality to evaluate which medications have a significant influence on how antidepressants and anxiolytics are prescribed. @eq-eigen-centrality outlines the calculation of eigenvector centrality; $c_i$ and $c_j$ are the centralities of node $\mathcal{v}_i$ and $\mathcal{v}_j$, respectively; $\lambda$ is the eigenvalue; $a_{ji}$ is an element on row $j$ and column $i$ from an adjacency matrix $A$, representing the connection between node $\mathcal{v}_j$ and $\mathcal{v}_i$.

$$
\displaystyle c_i = \frac{1}{\lambda} \sum_{j \neq i} a_{ji} \cdot c_j
$$ {#eq-eigen-centrality}

@@

Classical and seasonal-trend decomposition techniques may not fully capture complex periodic patterns in time-series data. Singular spectrum analysis (SSA) is a non-parametric method that leverages Hankel matrix embedding for decomposition. In this study, SSA was applied to identify trends and seasonal components in weekly aggregated prescription data. To better isolate trend and seasonal components, sequential SSA was applied. First, a basic SSA model with the lag parameter $L$ = 52 was used to extract the dominant trend. The residuals were then analyzed with a second SSA model using $L = \frac{N}{2}$ to capture complex periodic patterns. The extracted trend, residuals, and oscillatory components were evaluated visually and statistically, with Mann-Kendall trend tests applied to assess long-term changes. A complete documentation of this approach and its rationale was included in the supplementary file section S1.1.

### 2025-03-12 docs/exec-summary/_descriptive.qmd
Revise the methods section

The data was extracted as a daily time-series, which then aggregated into weekly and monthly for exploratory data analysis. The data contains five metrics, including the number of daily medication claims, the number of daily patients, claim-to-patient ratio, and eigenvector centrality. For each metrics in each dataset, univariate descriptive analysis was performed by calculating the statistical summary to describe the centrality and spread. Mean, median, standard deviation, and interquartile range were reported to evaluate the distribution.

>>>>

The data was extracted as a daily time series and aggregated into weekly and monthly intervals for analysis. It includes four metrics: daily medication claims, daily patients, claim-to-patient ratio, and eigenvector centrality. For each metric, a univariate descriptive analysis was performed, reporting the mean, median, standard deviation, and interquartile range to summarize centrality and dispersion.

### 2025-03-12 docs/exec-summary/_source-data.qmd
Add date of database access #23
Revise the methods section

All data in this investigation originated from IADB.nl, a dynamic pharmacy database containing prescription data from 1994 which still grow expansively. The database covers approximately 120 community pharmacies serving over one million patients. All patients are recorded in the database, irrespective of health care insurance, where the prescription rates, age, and gender are generalizable to represent the Netherlands [@Visser2013]. The database offers a longitudinal record, where each individual has a uniquely anonymized tracking identifier. All records contain fields on the date of medication dispensing, its quantity, dose regimen, number of days prescribed, prescribing physicians, and the Anatomical Therapeutic Chemical (ATC) code. Except for the over the counter drugs and medication dispensed during hospitalization, other medication records for each patient are complete, mainly due to high patient-pharmacy commitment in the Netherlands. This analysis include daily medication claims from a static cohort of adults aged 18-65 years old and prescribed for anxiolytics or antidepressants at least once in the time span from 2018 to 2022.

>>>>

All data in this study originated from IADB.nl, a dynamic pharmacy database containing prescription data from 1994 which still grow expansively. The database covers approximately 120 community pharmacies serving over one million patients and was accessed on the 20^th^of September, 2023. All patients are recorded in the database, irrespective of health care insurance, where the prescription rates, age, and gender are generalizable to represent the Netherlands [@Visser2013]. The database offers a longitudinal record, where each individual has a uniquely anonymized tracking identifier. All records contain fields on the date of medication dispensing, its quantity, dose regimen, number of days prescribed, prescribing physicians, and the Anatomical Therapeutic Chemical (ATC) code. Except for the over the counter drugs and medication dispensed during hospitalization, other medication records for each patient are complete, mainly due to high patient-pharmacy commitment in the Netherlands. This analysis includes daily medication claims from a static cohort of adults aged 18-65 years old and prescribed for anxiolytics or antidepressants at least once in the time span from 2018 to 2022.

### 2025-03-12 docs/exec-summary/_spectral.qmd
Revise the formal definition following the cited references
Revise the methods section

Classical and STL-based decomposition technique may not sufficiently capture complex periodic patterns in a time-series. Singular spectrum analysis (SSA) is a powerful non-parametric technique by leveraging Hankel matrix as a higher-dimensional embedding of the time-series. The resulting Hankel matrix $X$ has the size of $L \times K$, where $L$ represents the lag term. The lag term is the length of data point in a time-series being taken to construct the higher-dimensional embedding. For a weekly data with a hypothesized yearly seasonality, the $L$ is set as 52, representing the number of week in a year. In theory, the number of $L$ should be between the range of $2 \leq L \leq \frac{N}{2}$, where $N$ is the total length of the time-series [@Golyandina2020].

@@

The embedded time-series $X$ is then decomposed using a singular value decomposition (SVD). With a matrix $S = X \cdot X^T$, we can extract an eigenvalue $\lambda \ni \{\lambda_1 \geq \dots \geq \lambda_L\}$, where each $\lambda_i$ is a non-negative integer. Similarly, we can define $U \ni \{U_1, \dots, U_L\}$ as the eigenvectors of matrix $S$ corresponding to eigenvalues $\lambda$. Afterwards, we can extract factor vectors $V$ where the corresponding $V_i = X^T \frac{U_i}{\sqrt{\lambda_i}}$. Finally, the eigentriple of $\{\sqrt{\lambda_i}, U_i, V_i^T\}$ is formulated as a row-wise decomposition of matrix $X$ [@Golyandina2020]. The resulting eigentriple is then used to reconstruct the time-series by eigentriple grouping and diagonal averaging.

>>>>

Classical and Seasonal-Trend decomposition technique may not sufficiently capture complex periodic patterns in a time-series. Singular spectrum analysis (SSA) is a powerful non-parametric technique by leveraging Hankel matrix as a higher-dimensional embedding of the time-series. The resulting Hankel matrix $X$ has the size of $L \times K$, where $L$ represents the lag term. The lag term is the length of data point in a time-series being taken to construct the higher-dimensional embedding. For a weekly data with a hypothesized yearly seasonality, the $L$ is set as 52, representing the number of week in a year. In theory, the number of $L$ should be between the range of $2 \leq L \leq \frac{N}{2}$, where $N$ is the total length of the time-series [@Golyandina2020].

@@

The embedded time-series $X$ is then decomposed using a singular value decomposition (SVD). With a matrix $S = X \cdot X^T$, we can extract an eigenvalue $\lambda \ni \{\lambda_1 \geq \dots \geq \lambda_L\}$, where each $\lambda_i$ is a non-negative integer. Similarly, we can define $U \ni \{U_1, \dots, U_L\}$ as the eigenvectors of matrix $S$ corresponding to eigenvalues $\lambda$. Afterwards, we can extract factor vectors $V \ni \{V_1, \dots, V_L\}$ where the corresponding $V_i = X^T \frac{U_i}{\sqrt{\lambda_i}}$. Finally, the eigentriple of $\{\sqrt{\lambda_i}, U_i, V_i^T\}$ is formulated as a row-wise decomposition of matrix $X$ [@Golyandina2020]. The resulting eigentriple is then used to reconstruct the time-series by eigentriple grouping and diagonal averaging.
